Review of Edge Detectors

### 2014-11-18 edge-detection-review-3rd-revision-v2.tex
added 2nd version of review and missing files

% en este parrafo, puede ser cambiar trough por valley
In 1980 Marr and Hildreth \cite{segm:edge_region:marr:80:digital_step} argued that intensity changes are not independent 
of image scale, so edge detection requires the use of different size 
operators. They also argued that a sudden intensity change will be seen 
as a peak (or valley) in the first derivative or, equivalently, as a zero 
crossing in the second derivative. This algorithm is presented in Section \ref{sec:second}. 
Haralick's algorithm \cite{bb20239} is an alternative approach based on the second derivative 
which is also reviewed in Section \ref{sec:second}. This algorithm has the particularity of 
proposing a model to locally approximate the image around a point; then, using this model, 
an approximation to the second derivative of the image can be calculated analytically and 
finding edges is achieved by imposing a condition over the model parameters.

@@

The simplest digital approximation to the partial derivatives using masks of size $3\times3$ are obtained 
by taking the difference between the third and first rows (or columns) of the $3\times3$ region. The
difference between the third and first rows approximates the derivative in the x-direction, and 
the difference between the third and first columns approximates the derivative in the y-direction.
These approximations can be implemented by filtering the image with the two masks in Figure \ref{fig:prewitt}.
These masks are called \textit{Prewitt operators} \cite{prewitt1970object}.

@@

\myn{\label{second:references}}The two algorithms presented in this section (Marr-Hildreth \cite{segm:edge_region:marr:80:digital_step} and Haralick \cite{bb20239}) 
are based on the second derivative of the image, and both take steps to reduce noise before 
detecting edges in the image.

@@

Zero crossing detection is the key feature of the Marr-Hildreth edge detection method. The technique 
presented in the previous paragraph is attractive for its simplicity of implementation and its low 
computational cost. In general it is a technique that yields good results, but if more precision in 
finding the zero crossings is needed, more advanced methods for finding zero crossings with subpixel 
accuracy could be used (e.g.\ marching squares \cite{marching_cubes}).

@@

If we substitute the above expressions in the bi-cubic polynomia of eq., we obtain

@@

\begin{equation}
\exists\rho \in [0,\rho_0) \:\:  \mathtt{such\:that}  \:\:  2C_2 + 6C_3\rho = 0 \\
\end{equation}

@@

In Figure \ref{fig:results:parameters:roberts:saliency} we show a summary of the threshold's behavior for this algorithm, in the whole range of possible values $[0,1]$. For each image, this information is summarized in a saliency map which serves to asses the quality of the hierarchical set of edges obtained.
The saliency map is constructed by sampling the scale range (using steps of $0.05$) and computing the edge map corresponding for each scale. Then each edge is labeled with its scale of disappearance $\alpha^-$, which is the maximum scale at which an edge is present in the output edge map. This scale is drawn with a \emph{hot} colormap (black, red, yellow, white). Intuitively, this means that 'hotter' edges last longer across scales and thus are more meaningful.
These results show two important properties. First, the obtained stack of edges are causal, i.e. as the scale parameter increases the edge map gets coarser. This can be verified by observing the values of $\alpha^-$ assigned to each contour.
If a contour has a high $\alpha^-$ value it means that it will survive for a long time in the merging process. On the other hand, small values mean that it will be removed at early stages.
As the figure shows, the highest values of $\alpha^-$ are assigned to the most coarser segmentation.
And second, the stack of edges is geometrically accurate, i.e. the borders present in each scale are a strict subset of the previous (finer) scale. If they were not included they will be seen as new contours in the saliency map.
These results empirically show that the obtained hierarchy holds the strong causality property introduced by Morel and Solimini \cite{book:morel:95:variational_pde}.

@@

\renewcommand{\anchocinco}{2.5cm}
\newcommand{\anchocuatro}{5cm}

@@

\renewcommand{\anchocuatro}{5cm}

@@

\subsection{Discussion}
\myn{\label{results:discussion}}In this section we will discuss the obtained results and contrast them with the claims made by the original authors.
In the case of the first derivative algorithms, they are very simple and the claims made by the original authors are straight forward to verify. The Roberts algorithm was presented in his Phd Thesis (\cite{im_proc:segmentation:roberts:1963:boundary_extraction}, Chapter IV ) as a step of a more complex 3D reconstruction system. For this reason, only qualitative results are shown on few images. The effects of the parameter or the discretization is not validated. The only observation made in that work is the presence of spurious edges due to noise (pg. 31), which are removed by a line fitting postprocessing.
With respect to Prewitt and Sobel Operators,

In Haralick's work, there is an interesting quantitative validation over a synthetic grid. On that evaluation, sensitivity to noise as well as comparative performance with Prewitt operators. As in this work we didn't benchmark quantitaive

The obtained results validate the original claims made by Marr \cite{segm:edge_region:marr:80:digital_step}, were he made the following observations:
\begin{itemize}
\item "A major difficulty with natural images is that changes can and do occur over a
wide range of scales (Marr 1976a, b). No single filter can be optimal simultaneously at all scales, so it follows that one should seek a way of dealing separately with the changes occurring at different scales. ".

@@

The first claim is validated in our experiments varying the $\sigma$ parameter, where we observed that no single value will allow us to obtain all the meaningful edges. The same reasoning applies to the zero crossing threshold $t_{ZC}$ .

@@

With respect to the parameters, these methods require the manual determination of the parameters, and it is very hard to find a single threshold value suitable for a broad category of examples. In this sense, \emph{A contrario} methods provide a better approach for setting parameters.
For this reason, most modern approaches use a multiscale approach, were edges are detected at many different scales and the optimum scale is chosen a posteriori, as proposed in \cite{morphology:watershed:najman:1996:geodesic_saliency} or \cite{im_proc:segmentation:arbelaez:2006:boundary_extraction}.
This idea can be even extended to pick edges at different scales depending on the region of the image (see \cite{segm:cardelino:2012:a_contrario_selection_of_optimal_partitions} for a similar idea applied to regions).

@@

The goal of this work was to provide reference implementation for some of the most classical edge detectors. Some of them have been published for many years and are quite simple, however, it is surprisingly hard to find trustable implementations. We focused on low-level edge detection, which means that we do not consider post processing steps to integrate the detected edgels into meaningul lines.
Extensive qualitative validation has been performed, showing the effects of the parameters and the scale-space properties of the algorithms in order to faithfully compare them.
We consider this work as the first step towards a complete benchmark of edge detectors, which has to be completed with a systemic quantitative analysis. 
A suitable framework for this quantitative analysis is discussed and applied in \cite{segm:cardelino:2012:a_contrario_selection_of_optimal_partitions}, but this is clearly outside the scope of this work.
In particular, quantitative challenging the claims of the reviewed works will be a great contribution for the area, which is very weak with respect to experimental validation. From the original works reviewed, only Haralick presents a quantitative analysis but only over one synthetic images. Even at the moment of writing (2014), there are only a dozen edge detection and segmentation works presenting a serious quantitative analysis over a standard database of images.

>>>>

% en este parrafo, puede ser cambiar trough por valley
In 1980 Marr and Hildreth \cite{segm:edge_region:marr:84:digital_step} argued that intensity changes are not independent 
of image scale, so edge detection requires the use of different size 
operators. They also argued that a sudden intensity change will be seen 
as a peak (or valley) in the first derivative or, equivalently, as a zero 
crossing in the second derivative. This algorithm is presented in Section \ref{sec:second}. 
Haralick's algorithm \cite{bb20239} is an alternative approach based on the second derivative 
which is also reviewed in Section \ref{sec:second}. This algorithm has the particularity of 
proposing a model to locally approximate the image around a point; then, using this model, 
an approximation to the second derivative of the image can be calculated analytically and 
finding edges is achieved by imposing a condition over the model parameters.

@@

The simplest digital approximation to the partial derivatives using masks of size $3\times3$ are obtained 
by taking the difference between the third and first rows (or columns) of the $3\times3$ region. The
difference between the third and first rows approximates the derivative in the x-direction, and 
the difference between the third and first columns approximates the derivative in the y-direction.
These approximations can be implemented by filtering the image with the two masks in Figure \ref{fig:prewitt}.
These masks are called \textit{Prewitt operators} \cite{im_proc:segmentation:prewitt:1970:object_enhancement}.

@@

\myn{\label{second:references}}The two algorithms presented in this section (Marr-Hildreth \cite{segm:edge_region:marr:84:digital_step} and Haralick \cite{bb20239}) 
are based on the second derivative of the image, and both take steps to reduce noise before 
detecting edges in the image.

@@

Zero crossing detection is the key feature of the Marr-Hildreth edge detection method. The technique 
presented in the previous paragraph is attractive for its simplicity of implementation and its low 
computational cost. In general it is a technique that yields good results, but if more precision in 
finding the zero crossings is needed, more advanced methods for finding zero crossings with subpixel 
accuracy could be used (e.g.\ marching squares \cite{comp_graph:surf_recon:lorensen:87:marching_cubes}).

@@

If we substitute the above expressions in the bi-cubic polynomial of Eq. \ref{eq:bicubic}, we obtain

@@

\begin{equation}
\exists\rho \in (-\rho_0,\rho_0) \:\:  \mathtt{such\:that}  \:\:  2C_2 + 6C_3\rho = 0 \\
\end{equation}

@@

In Figure \ref{fig:results:parameters:roberts:saliency} we show a summary of the threshold's behavior for this algorithm, in the whole range of possible values $[0,1]$. For each image, this information is summarized in a saliency map which serves to assess the quality of the obtained hierarchical set of edges.
The saliency map is constructed by sampling the scale range (using steps of $0.05$) and computing the edge map corresponding for each scale. Then each edge is labeled with its scale of disappearance $\alpha^-$, which is the maximum scale at which an edge is present in the output edge map. This scale is drawn with a \emph{hot} colormap (black, red, yellow, white). Intuitively, this means that 'hotter' edges last longer across scales and thus are more meaningful.
These results show two important properties. First, the obtained stack of edges are causal, i.e. as the scale parameter increases the edge map gets coarser. This can be verified by observing the values of $\alpha^-$ assigned to each contour.
If a contour has a high $\alpha^-$ value it means that it will survive for a long time in the merging process. On the other hand, small values mean that it will be removed at early stages.
As the figure shows, the highest values of $\alpha^-$ are assigned to the most coarser segmentation.
And second, the stack of edges is geometrically accurate, i.e. the borders present in each scale are a strict subset of the previous (finer) scale. If they were not included they will be seen as new contours in the saliency map.
These results show empirical evidence that the obtained hierarchy holds the strong causality property, introduced by Morel and Solimini \cite{book:morel:95:variational_pde}.

@@

\renewcommand{\anchocinco}{2.5cm}
\newcommand{\anchocuatro}{4.5cm}

@@

\renewcommand{\anchocuatro}{4.5cm}

@@

\subsection{Discussion}
\myn{\label{results:discussion}}In this section we will discuss the obtained results and contrast them with the claims made by the original authors.
In the case of the first derivative algorithms, they are very simple and the claims made by the original authors are straight forward to verify. The Roberts algorithm was presented in his Phd Thesis (\cite{im_proc:segmentation:roberts:1963:boundary_extraction}, Chapter IV ) as a step of a more complex 3D reconstruction system. For this reason, only qualitative results are shown on few images. The effects of the parameter or the discretization is not validated. The only observation made in that work is the presence of spurious edges due to noise (pg. 31), which are removed by a line fitting postprocessing.
With respect to \emph{Prewitt} and \emph{Sobel} operators, they are arise from an in-depth study of different ways to discretize the gradient. In \cite{im_proc:segmentation:prewitt:1970:object_enhancement} these discretizations are discussed and compared qualitatively. In this work we tested only two of the possible discretizations, and we obtained very similar results using both operators. On the other hand, they show better performance that the \emph{Roberts} mask;

In Haralick's work, there is an interesting quantitative validation over a synthetic grid. On that evaluation, he shows results of noise sensitivity as well as comparative performance with Prewitt operators. As in this work we didn't focus on quantitative benchmarks, we didn't reproduce those results.

The obtained results validate the original claims made by Marr \cite{segm:edge_region:marr:84:digital_step}, where he made the following observations:
\begin{itemize}
\item "A major difficulty with natural images is that changes can and do occur over a
wide range of scales (Marr 1976a, b). No single filter can be optimal simultaneously at all scales, so it follows that one should seek a way of dealing separately with the changes occurring at different scales. ".

@@

The first claim is validated in our experiments by varying the $\sigma$ parameter, where we observed that no single value will allow us to obtain all the meaningful edges. The same reasoning applies to the zero crossing threshold $t_{ZC}$ .

@@

With respect to the parameters, these methods require the manual determination of the parameters, and it is very hard to find a single threshold value suitable for a broad category of examples. In this sense, \emph{A contrario} methods provide a better approach for setting parameters.
For this reason, most modern approaches use a multiscale approach, were edges are detected at many different scales and the optimum scale is chosen a posteriori, as proposed in \cite{morphology:watershed:najman:1996:geodesic_saliency} or \cite{im_proc:segmentation:arbelaez:2006:boundary_extraction}.
This idea can be even extended to pick edges at different scales depending on the region of the image (see \cite{segm:cardelino:2013:a_contrario_selection_of_optimal_partitions} for a similar idea applied to regions).

@@

The goal of this work was to provide reference implementations for some of the most classical edge detectors. Some of them have been published for many years and are quite simple, however, it is surprisingly hard to find trustable implementations. We focused on low-level edge detection, which means that we do not consider post processing steps to integrate the detected edgels into meaningul lines.
Extensive qualitative validation has been performed, showing the effects of the parameters and the scale-space properties of the algorithms in order to faithfully compare them.
We consider this work as the first step towards a complete benchmark of edge detectors, which has to be completed with a systemic quantitative analysis. 
A suitable framework for this quantitative analysis is discussed and applied in \cite{segm:cardelino:2013:a_contrario_selection_of_optimal_partitions}, but this is clearly outside the scope of this work.
In particular, quantitative challenging the claims of the reviewed works will be a great contribution for the area, which is very weak with respect to experimental validation. From the original works reviewed, only Haralick presents a quantitative analysis but only over one synthetic image. Even at the moment of writing (2014), there are only a dozen edge detection and segmentation works presenting a serious quantitative analysis over a standard database of images.
